---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

#load data

```{r}
candy_data <- read_csv(here::here("clean_data/candys_data_clean.csv"))
```
```{r}

#number of rating (not incluing nas)

candy_data %>% 
  drop_na(rating) %>% 
  summarise(n = n())


```

#average age of people who are going trick or treating

```{r}

#dropping nas
candy_data %>% 
  drop_na(age) %>% 
  filter(going_out == "Yes") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

#imputing with mean
mean_age_overall <- candy_data %>% 
  drop_na(age) %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age)) %>% 
  pull(avg_age)  
  
candy_data %>% 
  mutate(age = coalesce(age, mean_age_overall)) %>% 
  filter(going_out == "Yes") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

#imputing with median
median_age_overall <- candy_data %>%
  drop_na(age) %>% 
  group_by(id) %>%
  summarise(med_age = median(age)) %>% 
  summarise(med_age = median(med_age)) %>% 
  pull(med_age)

candy_data %>% 
  mutate(age = coalesce(age, median_age_overall)) %>% 
  filter(going_out == "Yes") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

```

#average age of people who are not going trick or treating

```{r}

#dropping nas
candy_data %>% 
  drop_na(age) %>% 
  filter(going_out == "No") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

#imputing with mean
  
candy_data %>% 
  mutate(age = coalesce(age, mean_age_overall)) %>% 
  filter(going_out == "No") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

#imputing with median

candy_data %>% 
  mutate(age = coalesce(age, median_age_overall)) %>% 
  filter(going_out == "No") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))


  

```


#For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
candy_data %>% 
  filter(rating == "JOY") %>% 
  group_by(candy) %>% 
  summarise(number_joy = n()) %>% 
  slice_max(number_joy, n = 3)

candy_data %>% 
  filter(rating == "DESPAIR") %>% 
  group_by(candy) %>% 
  summarise(number_despair = n()) %>% 
  slice_max(number_despair, n = 3)

candy_data %>% 
  filter(rating == "MEH") %>% 
  group_by(candy) %>% 
  summarise(number_meh = n()) %>% 
  slice_max(number_meh, n = 3)
```
#How many people rated Starburst as despair?


```{r}
candy_data %>% 
  filter(candy == "starburst" & rating == "DESPAIR") %>% 
  summarise(starburst_despair = n())
```


##For the next three questions, count despair as -1, joy as +1, and meh as 0.

#What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
candy_data %>%
  drop_na(rating) %>%
  mutate(rating_points = case_when(rating == "JOY" ~ 1,
                                   rating == "MEH" ~ 0,
                                   TRUE ~ -1)) %>% 
  group_by(candy) %>% 
  summarise(score = sum(rating_points)) %>% 
  slice_max(score, n = 3)
```


#What was the most popular candy bar in each year?

```{r}
candy_data %>%
  drop_na(rating) %>%
  mutate(rating_points = case_when(rating == "JOY" ~ 1,
                                   rating == "MEH" ~ 0,
                                   TRUE ~ -1)) %>% 
  group_by(candy, year) %>% 
  summarise(score = sum(rating_points)) %>% 
  ungroup() %>% 
  group_by(year) %>% 
  slice_max(score, n = 3)
```


#What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
candy_data %>%
  drop_na(rating, country) %>%
  mutate(
    rating_points = case_when(rating == "JOY" ~ 1,
                              rating == "MEH" ~ 0,
                              TRUE ~ -1),
    country = case_when(country == "usa" |
                        country == "uk" |
                        country == "canada" ~ country,
                        TRUE ~ "other country")
  ) %>% 
  group_by(country, candy) %>% 
  summarise(score = sum(rating_points)) %>% 
  slice_max(score, n = 1)

```



