---
title: "Candies Task"
output: 
  html_notebook:
    number_sections: yes

---

# Cleaning


## Looking at raw data

I looked through each of the data sets below noting the important columns, differences between the data sets and things I need to clean.

* column names to be made the same
* data needs to be pivoted to allow a join
* need to add a unique id for each person for grouping when answering future questions
* timestamp can be replaced with straight up year column


```{r}
candy_2015 <- read_excel(here::here("raw_data/boing-boing-candy-2015.xlsx"))
head(candy_2015)

candy_2016 <- read_excel(here::here("raw_data/boing-boing-candy-2016.xlsx"))
head(candy_2016)

candy_2017 <- read_excel(here::here("raw_data/boing-boing-candy-2017.xlsx"))
head(candy_2017)
```

## Pre bind cleaning

The below code shows:

* selection of important columns
* creation of year and id 
* also for 2015 country and gender columns needed to be created as they weren't present in the original data
* columns are renamed to be consistent with other tables
* columns are reordered
* table is pivoted put all the candy bar columns in to one long **candy** column and their ratings into a **rating** column
* candy column mutated to remove brackets at beginning and end

A similar process was followed for the 2016 and 2017 data
  
```{r}
candy_2015_piv <- candy_2015 %>%
  select(c(2:96)) %>%
  mutate(
    year = 2015,
    id = c(paste0(1:(nrow(
      candy_2015
    )), "_", 2015)),
    country = as.character(NA),
    gender = as.character(NA)
  ) %>%
  rename(
    age = "How old are you?",
    going_out = "Are you going actually going trick or treating yourself?",
    "[JoyJoy (Mit Iodine!)]" = "[JoyJoy (Mit Iodine)]"
  ) %>%
  select(id, year, going_out, gender, age, country, c(3:95)) %>%
  rename_all( ~ str_to_lower(.)) %>%
  pivot_longer(
    cols = c("[butterfinger]":"[york peppermint patties]"),
    names_to = "candy",
    values_to = "rating"
  ) %>%
  mutate(candy = str_sub(candy, 2, -2))
```


Output of the above code gives a really long table with 8 columns. Column names and number are consistent now across all three tables. 


```{r echo=TRUE}


head(candy_2015_piv)

```


```{r include=FALSE}
# hidden

candy_2016_piv <- candy_2016 %>%
  select(c(2:106)) %>%
  mutate(year = 2016,
         id = c(paste0(1:(nrow(
           candy_2016
         )), "_", 2016))) %>%
  select(id, year, c(1:4, 6:105)) %>%
  rename(
    going_out = "Are you going actually going trick or treating yourself?",
    gender = "Your gender:",
    age = "How old are you?",
    country = "Which country do you live in?",
    "[box’o’ raisins]" = "[Box'o'Raisins]"
  ) %>%
  rename_all( ~ str_to_lower(.)) %>%
  pivot_longer(
    cols = c("[100 grand bar]":"[york peppermint patties]"),
    names_to = "candy",
    values_to = "rating"
  ) %>%
  mutate(candy = str_sub(candy, 2, -2)) 


candy_2017_piv <- candy_2017 %>%
  select(c(2:109)) %>%
  mutate(year = 2017,
         id = c(paste0(1:(nrow(
           candy_2017
         )), "_", 2017))) %>%
  select(id, year, c(1:4, 6:108)) %>%
  rename_all( ~ str_replace(., "^Q[0-9]: ", "")) %>%
  rename_all( ~ str_to_lower(.)) %>%
  rename(
    "q6 | anonymous brown globs that come in black and orange wrappers" =
      "q6 | anonymous brown globs that come in black and orange wrappers	(a.k.a. mary janes)",
    "q6 | box’o’ raisins" = "q6 | box'o'raisins"
  )   %>%
  pivot_longer(
    cols = c("q6 | 100 grand bar":"q6 | york peppermint patties"),
    names_to = "candy",
    values_to = "rating"
  ) %>%
  mutate(candy = str_sub(candy, 6)) %>%
  clean_names()


```


## Bind and post bind cleaning 

* Using bind_rows the three tables were bound together.
* I then investigate the distinct values in the country.
* I grouped the distinct values into the below groups. 
* Any entry that reasonably be assumed to be reffering to a specific country was added to the group for that country i.e. Usa! Usa! Usa! assumed to be usa. 
* Countries that ony had one unique entry could be grouped together.
* All other entries assumed invalid and were set to NA
* Age column mutated to remove invalid entries.
* Invalid entries were taken as entries that were numeric characters (numbers with and without decimal points were included).
* Invalid entries set to 0.
* Type chnaged to numeric.
* Valid age range assumed to be  5 < age < 101, anything outside of this range set 0.
* All 0's set to NA

```{r}
candy_all_years <-
  bind_rows(candy_2015_piv, candy_2016_piv, candy_2017_piv) %>%
  mutate(
    country = tolower(country),
    country =
      case_when(
        country %in% usa_group ~ "usa",
        country %in% uk_group ~ "uk",
        country %in% spain_group ~ "spain",
        country %in% south_korea_group ~ "south korea",
        country %in% netherlands_group ~ "netherlands",
        country %in% canada_group ~ "canada",
        country %in% other_valid_countries ~ country,
        TRUE ~ "unknown"
      ),
    country = na_if(country, "unknown")
  ) %>%
  mutate(
    age = case_when(str_detect(age, "[^0-9\\.]") ~ "0",
                    TRUE ~ age),
    age = round(as.numeric(age)),
    age = case_when(age < 101 & age > 5 ~ age,
                    TRUE ~ 0),
    age = na_if(age, 0)
  )

```

## Write to csv

* New cleaned data written to csv file.


# Analysing Data

## Read in clean data

* clean data read into analysis notebook

## Answer question

Q1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

* There is a ratings column that contains all of the ratings made across all three year. 
* I just counted the number of rows.
* First I dropped NAs on rating so that the missing values were excluded.


```{r echo=TRUE}
candy_data %>% 
  drop_na(rating) %>% 
  summarise(n = n())
```

Q2. What was the average age of people who are going out trick or treating?

*  I used 3 different methods to solve this: droping NAs on age, imputing with mean and imputing with median.
* It is important to group by id here as the origninal columns had different numbers of question meaning after the pivot each persons age appears in a different number of rows. Grouping by id collapses this back down so that each persons age appears once, therefore not skewing the averages. 

```{r echo=TRUE}
#dropping nas
candy_data %>% 
  drop_na(age) %>% 
  filter(going_out == "Yes") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

#imputing with mean
mean_age_overall <- candy_data %>% 
  drop_na(age) %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age)) %>% 
  pull(avg_age)  
  
candy_data %>% 
  mutate(age = coalesce(age, mean_age_overall)) %>% 
  filter(going_out == "Yes") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

#imputing with median
median_age_overall <- candy_data %>%
  drop_na(age) %>% 
  group_by(id) %>%
  summarise(med_age = median(age)) %>% 
  summarise(med_age = median(med_age)) %>% 
  pull(med_age)

candy_data %>% 
  mutate(age = coalesce(age, median_age_overall)) %>% 
  filter(going_out == "Yes") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))
```


Q2. What was the average age of people who are not going out trick or treating?

* Same as above but for people not going out. 

```{r echo=FALSE}

#dropping nas
candy_data %>% 
  drop_na(age) %>% 
  filter(going_out == "No") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

#imputing with mean
  
candy_data %>% 
  mutate(age = coalesce(age, mean_age_overall)) %>% 
  filter(going_out == "No") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))

#imputing with median

candy_data %>% 
  mutate(age = coalesce(age, median_age_overall)) %>% 
  filter(going_out == "No") %>% 
  group_by(id) %>% 
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age = mean(avg_age))


  

```


Q4. For each of joy, despair and meh, which candy bar received the most of these ratings?

* Filter by the rating 
* group by candy
* count the total number of the ratings for each rating
* slice of the top 3 for each rating

```{r echo=TRUE}
candy_data %>% 
  filter(rating == "JOY") %>% 
  group_by(candy) %>% 
  summarise(number_joy = n()) %>% 
  slice_max(number_joy, n = 3)

candy_data %>% 
  filter(rating == "DESPAIR") %>% 
  group_by(candy) %>% 
  summarise(number_despair = n()) %>% 
  slice_max(number_despair, n = 3)

candy_data %>% 
  filter(rating == "MEH") %>% 
  group_by(candy) %>% 
  summarise(number_meh = n()) %>% 
  slice_max(number_meh, n = 3)
```

Q5. How many people rated Starburst as despair?

* filter for starburst candy were the rating is despair
* count the number 

```{r echo=TRUE}
candy_data %>% 
  filter(candy == "starburst" & rating == "DESPAIR") %>% 
  summarise(starburst_despair = n())
```


**For the next three questions, count despair as -1, joy as +1, and meh as 0.**

6. What was the most popular candy bar by this rating system for each gender in the dataset?

* drop the na ratings
* create a new column with the points system
* group by candy
* the summaries sums the points for each candy and returns the total
* slice off the top three and show

```{r echo=TRUE}
candy_data %>%
  drop_na(rating) %>%
  mutate(rating_points = case_when(rating == "JOY" ~ 1,
                                   rating == "MEH" ~ 0,
                                   TRUE ~ -1)) %>% 
  group_by(candy) %>% 
  summarise(score = sum(rating_points)) %>% 
  slice_max(score, n = 3)
```


Q7. What was the most popular candy bar in each year?

* same as before
* group by year then candy
* summaries gives an entry for each candy bar for all three years individually.
* slice the top three from each year

```{r}
candy_data %>%
  drop_na(rating) %>%
  mutate(rating_points = case_when(rating == "JOY" ~ 1,
                                   rating == "MEH" ~ 0,
                                   TRUE ~ -1)) %>% 
  group_by(year, candy) %>% 
  summarise(score = sum(rating_points)) %>% 
  slice_max(score, n = 3)
```

Q8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

* Same as before but drop NAs on country as well as th NAs could be US, UK or Canda get put them into the other countries group.
* Mutate country column so that if country isn't usa, uk or canada the value is changed to **other country**.
* grouped by country and candy and summed as before.
* Sliced off top candy in each country


```{r}
candy_data %>%
  drop_na(rating, country) %>%
  mutate(
    rating_points = case_when(rating == "JOY" ~ 1,
                              rating == "MEH" ~ 0,
                              TRUE ~ -1),
    country = case_when(country == "usa" |
                        country == "uk" |
                        country == "canada" ~ country,
                        TRUE ~ "other country")
  ) %>% 
  group_by(country, candy) %>% 
  summarise(score = sum(rating_points)) %>% 
  slice_max(score, n = 1)
```

