---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(stringr)
library(janitor) 
```

#read in clean data

```{r}
clean_decathlon_results <- read_csv(here::here("task1_decathlon_data/clean_data/decatlon_clean_data.csv"))
```

#Who had the longest long jump seen in the data?

```{r}
clean_decathlon_results %>%
  select(name, event, event_performance) %>% 
  filter(event == "long_jump") %>% 
  arrange(desc(event_performance)) %>% 
  head(1)



```

#What was the average 100m time in each competition?

```{r}
clean_decathlon_results %>%
  select(competition, event, event_performance) %>% 
  filter(event == "100m") %>% 
  group_by(competition) %>% 
  summarise(avg_100m_time = mean(event_performance))

```

#Who had the highest total points across both competitions?

```{r}
# highest total across both competition - assuming that means sum if a competitor competed at both

clean_decathlon_results %>%
  select(name, competition, points) %>% 
  group_by(name, competition) %>% 
  summarise(points = max(points)) %>% 
  summarise(points = sum(points)) %>% 
  slice_max(points)
 

```

# What was the shot-put scores for the top three competitors in each competition?

```{r}
clean_decathlon_results %>% 
  filter(event == "shot_put") %>% 
  group_by(competition) %>% 
  slice_max(event_performance, n = 3)

```


# What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
clean_decathlon_results %>% 
  select(points, event, event_performance) %>% 
  filter(event == "400m") %>% 
  mutate(faster_than_50s = event_performance < 50) %>% 
  group_by(faster_than_50s) %>% 
  summarise(avg_points = mean(points))
```
