---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


Dog questions

```{r}
dog_data <- read_csv(here::here("clean_data/clean_dog_data.csv"))
```


Q1. The client only counts a valid email address as one ending in ‘.com’. How many survey results have a valid email address.

```{r}
dog_data %>% 
  filter(str_detect(email, "\\.com$")) %>% 
  summarise(valid_emails = n())
```


Q2. What’s the average amount spent on dog food for each dog size.

```{r}
dog_data %>% 
  drop_na(dog_size) %>% 
  group_by(dog_size) %>% 
  summarise(avg_amount_spent = mean(amount_spent_on_dog_food, na.rm = TRUE))
```


Q3. For owners whose surname starts with a letter in the second half of the alphabet (N onwards) what is the average age of their dog?

```{r}
dog_data %>% 
  filter(str_detect(last_name, "(?i)^[N-Z]")) %>% 
  summarise(avg_age_n_z = mean(dog_age, na.rm = TRUE))
```


Q4. The dog_age column is the age in dog years. If the conversion is 1 human year = 6 dog years, then what is the average human age for dogs of each gender?

```{r}
dog_data %>% 
  mutate(human_age = dog_age / 6) %>%
  drop_na(dog_gender) %>% 
  group_by(dog_gender) %>% 
  summarise(avg_human_age = mean(human_age, na.rm = TRUE))
```


Q5. Create a plot of results of question 4.

```{r}
counts <- dog_data %>% 
  mutate(human_age = dog_age / 6) %>%
  drop_na(dog_gender) %>% 
  group_by(dog_gender) %>% 
  summarise(avg_human_age = mean(human_age, na.rm = TRUE)) %>% 
  pull(avg_human_age)

names <- dog_data %>% 
  mutate(human_age = dog_age / 6) %>%
  drop_na(dog_gender) %>% 
  group_by(dog_gender) %>% 
  summarise(avg_human_age = mean(human_age, na.rm = TRUE)) %>% 
  pull(dog_gender)

barplot(height = counts, ylim = c(0, 10), names.arg = names, 
        ylab = "avg age in human years", xlab = "dog gender",
        main = "Avg Age vs Gender", )

```

















